---
title: "Activity monitoring data"
output: html_document
date: "2022-10-06"
---

Set the R environment
```{r}
library(lattice)
library("knitr")
opts_chunk$set(echo=TRUE, results="hold")
Sys.setlocale("LC_TIME", "English")
```

Loading and preprocessing the data
```{r}
dtRAW <- read.csv('activity.csv', header = TRUE, na.strings = "NA", stringsAsFactors=FALSE)
str(dtRAW)
head(dtRAW)
tail(dtRAW)
```

What is mean total number of steps taken per day?
```{r}
totalSPD <- aggregate(steps~date, data=dtRAW, FUN=sum, na.rm=TRUE)
str(totalSPD)
head(totalSPD)
tail(totalSPD)
```

```{r}
hist(totalSPD$steps, breaks=20, col="grey",
     main="Total number of steps taken per day \n (missing data removed)",
     xlab="Number of steps")
```

```{r}
muTSPD <- mean(totalSPD$steps)
meTSPD <- median(totalSPD$steps)
print(muTSPD)
print(meTSPD)
```

What is the average daily activity pattern?
```{r}
muSPI <- aggregate(steps~interval, data=dtRAW, FUN=mean, na.rm=TRUE)
plot(muSPI, type="l",
     main="Average number of steps per 5-minute interval",
     xlab="5-minute intervals",
     ylab="Average numberof steps")
```

```{r}
numMaxStep <- muSPI[which(muSPI$steps==max(muSPI$steps)), ]$interval
print(numMaxStep)
```

Imputing missing values
```{r}
numNA <- nrow(dtRAW[which(is.na(dtRAW$steps)),])
print(numNA)
```

The strategy for filling missing values will be replacing missing values with the average number of steps in each 5-minute interval.

```{r}
dtNEW <- transform(dtRAW,
                   steps=ifelse(is.na(dtRAW$steps),
                                muSPI[match(muSPI$interval, dtRAW$interval), ]$steps,
                                dtRAW$steps))
sum(is.na(dtNEW))
str(dtNEW)
head(dtNEW)
tail(dtNEW)
```

```{r}
newTSPD <- aggregate(steps~date, data=dtNEW, FUN=sum, na.rm=TRUE)
muNTSPD <- mean(newTSPD$steps)
meNTSPD <- median(newTSPD$steps)
hist(newTSPD$steps, breaks=20, col="grey",
     main="Total number of steps taken per day \n (missing data filled in)",
     xlab="Number of steps")
```

```{r}
print(muNTSPD)
print(meNTSPD)
```

```{r}
par(mfrow=c(1, 2))
hist(totalSPD$steps, breaks=20, col="grey",
     main="Total number of steps taken per day \n (missing data removed)",
     xlab="Number of steps")
hist(newTSPD$steps, breaks=20, col="grey",
     main="Total number of steps taken per day \n (missing data filled in)",
     xlab="Number of steps")
```

Are there differences in activity patterns between weekdays and weekends?
```{r}
dtNEW$date <- as.Date(dtNEW$date, "%Y-%m-%d")
dtNEW$weekdays <- as.factor(ifelse(weekdays(dtNEW$date) %in% c("Saturday", "Sunday"),
                                  "weekend", "weekday"))
str(dtNEW)
head(dtNEW)
tail(dtNEW)
```

```{r}
weekSPI <- aggregate(steps ~ weekdays + interval, dtNEW, FUN=mean, na.rm=TRUE)
xyplot(steps ~ interval | weekdays, data=weekSPI, type="l", layout=c(1, 2),
       xlab="Interval", ylab="Number of steps")
```
